---
title: "Week01_Weekend_Homework"
output: html_document
date: "2023-02-24"
---

As this is your first weekend homework, here are some tips: 

* Try to schedule some time in your weekend to work on the homework so it's not suddenly Monday morning and you haven't gotten started yet (it happens).
* Remember that the weekend homework is for your learning, so try to use it as an opportunity to apply and consolidate everything you've learned in the week.
* Also use it as an opportunity to spend a bit more time making your code readable and reproducible, by practising commenting and writing some text around your steps and findings. You will thank yourself later! 
  * This will be especially useful for this specific weekend homework as it's very open-ended and you will eventually forget your own thought process
* A bit obvious, but don't spend your entire weekend working on the homework! Remember to spend time doing things you enjoy and rest up ahead of the following week.

The data for this weekend homework is scraped from Goodreads (a website all about books) and made publicly available on Kaggle. You can read more about the data [here](https://www.kaggle.com/jealousleopard/goodreadsbooks).

# MVP

### First steps

Load necessary packages and read in `books.csv`. Investigate dimensions, variables, missing values - you know the drill!

### Up to you

Now it's up to you... For this weekend homework there will be no specific tasks, just you and this dataset! Using everything you've learned this week, try to describe/summarise at least 5 things about this dataset - using R and the tidyverse of course! Feel free to find and use new functions if there is something that the tidyverse doesn't offer, but do try to use this homework to apply what you have learned this week. Be prepared to share one of your findings on Monday!

### Remember

Before you submit, go through your weekend homework and make sure your code is following best practices as laid out in the `coding_best_practice` lesson.


```{r}
library(tidyverse)
```


### To retrieve your data (remember to leave data file at same level on path as the project- create project for the day first then save and then create a markdown or notebook file for the day on the same level)



```{r}
books <- read_csv("data/books.csv")
```
###Number of rows


```{r}
nrow(books)
```


####Number of columns 

```{r}
ncol(books)
```


#### Overall dimensions

```{r}
dim(books)
```

#### Variable names

```{r}
names(books)
```



#### Check the first 10 rows 

```{r}
head(books)
```



#### Check the last 10 rows 

```{r}
tail(books)
```


####Get an overview of the data with glimpse


```{r}
glimpse(books)
```



#### Look at the actual data in the table format > table appear top right next to the file 


```{r}
view(books)
```



##### To check if there are any missing NA values in the dataframe

```{r}
sum(is.na(books))
```

##### Select function

- Find the books with the highest average rating while keeping author, title, average rating and ratings count.


 [1] "rowid"              "bookID"             "title"              "authors"           
 [5] "average_rating"     "isbn"               "isbn13"             "language_code"     
 [9] "num_pages"          "ratings_count"      "text_reviews_count" "publication_date"  
[13] "publisher"         




```{r}
books %>% 
  select("title","authors","average_rating","ratings_count")
```

### to keep the top 5 highest ratings

```{r}
books %>% 
  select("title","authors","average_rating","ratings_count") %>% 
  arrange("average_rating") %>% 
  head()

```


### Rows with maximum values of a variable

```{r}
books %>% 
  #select("title","authors","average_rating","ratings_count") %>% 
  #arrange("average_rating") %>% 
  slice_max(isbn, n=5)

```



#### Filtered author.



```{r}
books %>% 
  filter(authors == "Tara MacCarthy")
```

### Filter all books with rating of 5

```{r}
books %>% 
  filter(average_rating == 5)
```




#### True or False Statement.

```{r}
1 == 3
  
```


#### True or False

## Works down the table row by row and will compare and take the rows that are True i.e average_rating is equal to 5


```{r}
books %>% 
  filter(average_rating == 5)
```


## True or False with higher > and lower than <

```{r}
books %>% 
  filter(average_rating > 4.9)
```



#### To find the one row that between 4.9 and 5


```{r}
books %>% 
  filter(average_rating > 4.9 & average_rating <5)
```




## Test if two things are not equal. True will bring non equals

```{r}
books %>% 
  filter(average_rating != 5)
```



#### Filter using & where it meets both conditions. If the average rating is superior or equal to 4 & the ratings_count is superior or equal to 200000




```{r}
books %>% 
  filter(average_rating >= 4 & ratings_count >=2000000)
```


#### Using or

```{r}
books %>% 
  filter(average_rating >= 4 | ratings_count >=2000000)
```



###Arrange title (arrange in ascending order > alphabetical )

```{r}
books %>% 
  arrange(title)
```



###Arrange in desceding order (counter wise)

```{r}
books %>% 
  arrange(desc(title))
```





#####

Takes first parameterand then arranges the second parameter 

So arranges average rating from 5 to 4 and inside there it will arrange alphabetically ascending first in the 5 then in the 4




```{r}
books %>% 
  arrange(desc(average_rating),tile))
```





### to arrange both in descending order needs to be done separately 

```{r}
books %>% 
  arrange(desc(average_rating), desc(title))
```





####mutate  > used to add new column.

Make total average rating column that creates the column .after = average_rating(can also write before)



```{r}
books %>% 
  mutate(total_average_rating =mean(average), .after = average_rating)
```




### filtered and arragned in descending order. 



```{r}
books %>% 
  mutate(total_average_rating = mean(average_rating), .after = average_rating) %>% 
  filter(average_rating > total_average_rating) %>% 
  arrange(desc(average_rating))
```



##### Create new total average column, made TRUE for every row which is higher than the average rathing and FALSE for every row that is lower that the average. 


```{r}
books %>% 
  mutate(total_average_rating = mean(average_rating), .after = average_rating) %>% 
   mutate(higher_than_average = (average_rating > total_average_rating))

```













####Group by > group all the books by same author are grouped together and then arrange by 

```{r}
books %>% 
  group_by(authors)
```



#### Summarise 

Summarise gives you the amount of entries in each group then the number of books by each author and the top 5 authors by numbers of books.


```{r}
books %>% 
  group_by(authors) %>% 
summarise(number_of_books = n()) %>% 
slice_max(number_of_books, n=5)
```



### Summarize

Summarise gives you the amount of entries in each group by authors and publisher then the number of books by each author and the top 5 authors by numbers of books.



```{r}

books %>%
  mutate(total_average_rating = mean(average_rating), .after = average_rating) %>%
   mutate(higher_than_average = (average_rating > total_average_rating), .after = total_average_rating) %>%
  group_by(higher_than_average, authors) %>%
summarise(number_of_books = n()) %>% 
  slice_max(number_of_books, n=5)



```




##### Mutate can change an existing column, takes information a different column to then make a new one and can just add a new column 

That makes new column with no data


```{r}
books %>% 
  mutate(colour_of_book = "")
```


### fills with na

```{r}

books %>% 
  mutate(colour_of_book = "na")
```


#### Assign numbers


```{r}
books %>% 
  mutate(colour_of_book = "1")
```




### Mutate by taking information and assigning it to a new column 


### too long is new column which is being placed after text_reviews_count and takes information from number of pages column and TRUE OR FALSE when compared to over 200 pages.

```{r}
books %>% 
  mutate(Too_long = (num_pages >200),.after = text_reviews_count)
```




##### Pull


```{r}
books %>% 
 mutate(average_number_of_pages = mean(num_pages)) %>% 
  summarise(avp = mean(num_pages)) %>% 
  pull(avp)

  
```





